---
title: "Full grid run"
author: "Tim Szewczyk"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


This is a wrapper for running the glossy buckthorn cellular automata model. It sets up the environment, allows the assignment of all necessary parameters, runs the simulation, and visualizes the results.

The land cover within each cell is compositional rather than the more common hard classification. As a result, all land cover-specific parameters are computed proportionally to the land cover in a cell (i.e., lc.prop %*% K, where lc.prop is a matrix with columns for land cover and a row for each cell, and K is a vector of carrying capacities with one per land cover type).

```{r wd.setup, include=FALSE}
if(!require(pacman)) {install.packages("pacman"); require(pacman)}
p_load("tidyverse", "magrittr", "stringr", "parallel", "here",
       "pbapply", "fastmatch", "gganimate"); theme_set(theme_bw())
library(gbPopMod)
data(lc.rct)
### lc.rct was produced from the raw GRANIT output as follows:
# grnt <- read_csv(here::here("data", "grnt_all.csv")) %>%
#   mutate(x=as.numeric(as.factor(.$left)),
#          y=as.numeric(factor(.$top, levels=rev(levels(factor(.$top))))),
#          x_y=paste(x, y, sep="_"))
# lc.rct <- as.tibble(expand.grid(x=1:max(grnt$x),
#                                 y=1:max(grnt$y))) %>%
#   mutate(x_y=paste(x, y, sep="_")) %>%
#   mutate(OpI=grnt$OpI[match(.$x_y, grnt$x_y)], 
#          Oth=grnt$Oth[match(.$x_y, grnt$x_y)], 
#          Dec=grnt$Dec[match(.$x_y, grnt$x_y)], 
#          Evg=grnt$Evg[match(.$x_y, grnt$x_y)], 
#          WP=grnt$WP[match(.$x_y, grnt$x_y)], 
#          Mxd=grnt$Mxd[match(.$x_y, grnt$x_y)],
#          inbd=!is.na(match(.$x_y, grnt$x_y)))
# lc.rct[is.na(lc.rct)] <- 0
```


```{r setPars}
##---
## set parameters
##---

  set.seed(225)
  n.sim <- 1
  g.p <- list(
    # general
    tmax=50,  # num time steps per simulation
    dem.st=FALSE,  # include stochasticity in demography?
    sdd.st=TRUE,  # include stochasticity in SDD?
    n_cores=4,  # parallelize sdd.pr calculation
    
    # landscape
    lc.r=100,  # num rows in landscape
    lc.c=100,  # num columns in landscape
    n.lc=6,  # num land cover categories
    N.p.t0=10,  # num cells with buckthorn at t=1
    
    # demography
    K=c(750, 10, 100, 100, 300, 100),  # K for adults
    pr.s=c(0.9, 0.1, 0.6, 0.6, 0.6, 0.6),  # pre-adult survival
    pr.f=c(0.9, 0.1, 0.29, 0.23, 0.2, 0.3),  # pr(fruit)
    fec=c(200, 100, 40, 20, 20, 10),  # mean(fruit per adult)
    age.f=3,#rep(c(3,5), c(2,4)),  # age @ first fruiting (length==n.lc | 1)
    bank=TRUE,  # include seedbank?
    pr.sb=0.3,  # pr(ann.surv seed bank)
    pr.est=c(0.07, 0.01, 0.08, 0.02, 0.02, 0.03),  # pr(seedling est)
    
    # dispersal
    sdd.max=15,  # max dispersal distance (unit: cells)
    sdd.rate=0.1,  # 1/mn for dispersal kernel
    pr.eat=c(0.3, 0.1, 0.2, 0.2, 0.2, 0.1),  # pr(birds eat fruit)
    bird.hab=c(.35, .35, 0.05, 0.1, 0.1, 0.05),  # bird habitat prefs
    pr.s.bird=0.6,  # seed viability post-digestion
    n.ldd=1   # num long distance dispersal events per year
  )
  
  control.p <- list(
    nTrt_grd=0.05,  # % of cells treated in each time step
    nTrt_man=0.05,  # % of cells treated in each time step
    grd.trt=c(Lit=0.005, Cov=0.01, Com=0.00001),
    man.trt=c(M=0.1, C=0.3, MC=0.8),
    add.owners=FALSE,  # do owners treat every year once starting?
    t.trt=20
  )

  

##--
## initialize landscape
##--
  
  # land cover
  lc.df <- lc.rct %>% 
    filter(y >= (max(lc.rct$y) - g.p$lc.r) & x <= g.p$lc.c) %>%
    mutate(id=row_number(), 
           id_inbd=min_rank(na_if(inbd*id, 0)))
  ngrid <- nrow(lc.df)
  ncell <- sum(lc.df$inbd)
    
  
  # populations
  # indexed 1:ngrid including rows for out-of-bounds cells
  N.init <- pop_init(ngrid, g.p, lc.df)
  
  # dispersal probabilities
  # indexed 1:ncell with only inbound cells: sdd.pr[,,,i] = ith inbound cell
  # row i = dplyr::filter(lc.df, inbds)[i,] = lc.df[id_inbd==i,]
  # cell indexes in sdd.pr[,,2,] correspond with lc.df$id
  # this extra complication avoids calculating many, many unnecessary cells
  system.time(sdd.pr <- sdd_set_probs(ncell, lc.df, g.p))
```

```{r runSim}
system.time(
for(s in 1:n.sim) {
  out <- run_sim(ngrid, ncell, g.p, lc.df, sdd.pr, N.init, control.p)
}
)
```


```{r plots, fig.height=8, fig.width=10}
out.sb <- log(out$N.sb+1)
out.ad <- out$N[,,max(g.p$age.f)]
N.out <- cbind(lc.df, out.ad) %>% as.tibble %>%
  gather(year, N.adult, (1:ncol(out.ad)) + ncol(lc.df)) %>%
  mutate(N.sb=c(out.sb),
         N.less5=c(out.sb > 0))
N.out$N.less5 <- N.out$N.less5 + c(out.ad > 5)
N.out$year <- str_pad(N.out$year, 3, "left", "0")

# final time step
adult.ab.fin <- ggplot(filter(N.out, year==str_pad(g.p$tmax+1, 3, "left", "0")),
                       aes(x=x, y=-y, fill=N.adult)) +
  geom_tile() + scale_fill_gradient(low="white", high="red") +
  ggtitle(paste("Adult abundance. Year", g.p$tmax+1))
sb.ab.fin <- ggplot(filter(N.out, year==str_pad(g.p$tmax+1, 3, "left", "0")), 
                    aes(x=x, y=-y, fill=N.sb)) +
  geom_tile() + scale_fill_gradient(low="white", high="red") +
  ggtitle(paste("log seed abundance. Year", g.p$tmax+1))
adult.pa.fin <- ggplot(filter(N.out, year==str_pad(g.p$tmax+1, 3, "left", "0")), 
                       aes(x=x, y=-y, fill=N.adult>0)) +
  geom_tile() + scale_fill_manual(values=c("FALSE"="white", "TRUE"="red")) + 
  ggtitle(paste("Adult presence. Year", g.p$tmax+1))
sb.pa.fin <- ggplot(filter(N.out, year==str_pad(g.p$tmax+1, 3, "left", "0")), 
                    aes(x=x, y=-y, fill=N.sb>0)) +
  geom_tile() + scale_fill_manual(values=c("FALSE"="white", "TRUE"="red")) +
  ggtitle(paste("Seed presence. Year", g.p$tmax+1))

# gifs
gifwd <- paste0("out/", ncell, "_t", g.p$tmax, "/")
if(!dir.exists(here::here("out"))) dir.create(here::here("out"))
if(!dir.exists(here::here(gifwd))) dir.create(here::here(gifwd))
age.i <- ifelse(length(g.p$age.f)==1, 
                paste0("age", g.p$age.f, ""),
                paste0("ages", min(g.p$age.f), "-", max(g.p$age.f), ""))
adult.ab <- ggplot(N.out, aes(x=x, y=-y, fill=N.adult, frame=year)) + 
  geom_tile() + scale_fill_gradient(low="white", high="red") + 
  ggtitle("Adult abundance. Year")
sb.ab <- ggplot(N.out, aes(x=x, y=-y, fill=N.sb, frame=year)) + 
  geom_tile() + scale_fill_gradient(low="white", high="red") + 
  ggtitle("log seed abundance. Year")
adult.pa <- ggplot(N.out, aes(x=x, y=-y, fill=N.adult>0, frame=year)) + 
  geom_tile() + scale_fill_manual(values=c("white","red")) + 
  ggtitle("Adult presence. Year")
sb.pa <- ggplot(N.out, aes(x=x, y=-y, fill=N.sb > 0, frame=year)) + 
  geom_tile() + scale_fill_manual(values=c("white","red")) + 
  ggtitle("Seed presence. Year")
adult.lo <- ggplot(N.out, aes(x=x, y=-y, fill=N.less5, frame=year)) +
  geom_tile() + ggtitle("Blue: Adult density ≤ 5. Year") + 
  scale_fill_gradient2(low="white", mid="blue", high="pink", midpoint=1)

# save plots
ggsave(paste0(gifwd, "Final_Ab_", age.i, ".jpg"), adult.ab.fin,
       width=8, height=6, units="in")
ggsave(paste0(gifwd, "Final_SB_Ab_", age.i, ".jpg"), sb.ab.fin,
       width=8, height=6, units="in")
ggsave(paste0(gifwd, "Final_PA_", age.i, ".jpg"), adult.pa.fin,
       width=8, height=6, units="in")
ggsave(paste0(gifwd, "Final_SB_PA_", age.i, ".jpg"), sb.pa.fin,
       width=8, height=6, units="in")
gganimate(adult.ab, paste0(gifwd, "Ab_", age.i, ".gif"), 
          interval=0.2, ani.width=800, ani.height=600)
gganimate(sb.ab, paste0(gifwd, "SB_Ab_", age.i, ".gif"),
          interval=0.2, ani.width=800, ani.height=600)
gganimate(adult.pa, paste0(gifwd, "PA_", age.i, ".gif"),
          interval=0.2, ani.width=800, ani.height=600)
gganimate(sb.pa, paste0(gifwd, "SB_PA_", age.i, ".gif"),
          interval=0.2, ani.width=800, ani.height=600)
gganimate(adult.lo, paste0(gifwd, "LoDens_", age.i, ".gif"),
          interval=0.2, ani.width=800, ani.height=600)


# # land cover
p.opi <- ggplot(N.out, aes(x=x, y=-y, fill=OpI)) + geom_tile() +
  scale_fill_gradient(low="white", high="red", limits=c(0,1))
ggsave(paste0(gifwd, "LC_OpI.jpg"), p.opi, width=10, height=7)

p.oth <- ggplot(N.out, aes(x=x, y=-y, fill=Oth)) + geom_tile() +
  scale_fill_gradient(low="white", high="gray30", limits=c(0,1))
ggsave(paste0(gifwd, "LC_Oth.jpg"), p.oth, width=10, height=7)

p.dec <- ggplot(N.out, aes(x=x, y=-y, fill=Dec)) + geom_tile() +
  scale_fill_gradient(low="white", high="green3", limits=c(0,1))
ggsave(paste0(gifwd, "LC_Dec.jpg"), p.dec, width=10, height=7)

p.wp <- ggplot(N.out, aes(x=x, y=-y, fill=WP)) + geom_tile() +
  scale_fill_gradient(low="white", high="orchid", limits=c(0,1))
ggsave(paste0(gifwd, "LC_WP.jpg"), p.wp, width=10, height=7)

p.evg <- ggplot(N.out, aes(x=x, y=-y, fill=Evg)) + geom_tile() +
  scale_fill_gradient(low="white", high="darkgreen", limits=c(0,1))
ggsave(paste0(gifwd, "LC_Evg.jpg"), p.evg, width=10, height=7)

p.mxd <- ggplot(N.out, aes(x=x, y=-y, fill=Mxd)) + geom_tile() +
  scale_fill_gradient(low="white", high="yellowgreen", limits=c(0,1))
ggsave(paste0(gifwd, "LC_Mxd.jpg"), p.mxd, width=10, height=7)
# 
# 
# # buckthorn vs land cover
# ggplot(N.out, aes(x=Dev)) + 
#   geom_point(aes_string(y=paste0("t_", g.p$tmax)), alpha=0.1)
# ggplot(N.out, aes(x=Oth)) + 
#   geom_point(aes_string(y=paste0("t_", g.p$tmax)), alpha=0.1)
# ggplot(N.out, aes(x=Hwd)) + 
#   geom_point(aes_string(y=paste0("t_", g.p$tmax)), alpha=0.1)
# ggplot(N.out, aes(x=WP)) + 
#   geom_point(aes_string(y=paste0("t_", g.p$tmax)), alpha=0.1)
# ggplot(N.out, aes(x=Evg)) + 
#   geom_point(aes_string(y=paste0("t_", g.p$tmax)), alpha=0.1)
# ggplot(N.out, aes(x=Mxd)) + 
#   geom_point(aes_string(y=paste0("t_", g.p$tmax)), alpha=0.1)
```





